Humann2 bar plot by myself

### 2019-10-17 scripts/09_barplot_myself.Rmd
update on the freaking bar plot

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

# the stratified table is used cuz we wanna know the species info
stratified <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified.tsv') %>% 
  rename(pathway = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$','')))

@@

### Starting from toxicity

```{r}
# get the pheno lable of the samples
phelabel <-  pheno %>% 
  select(Sampleid, Toxicity)

cnt <- stratified %>% 
  filter(str_detect(pathway, 'PWY-4242: pantothenate and coenzyme A biosynthesis III')) %>% 
  mutate(pathway = str_replace(pathway, 'PWY-4242: pantothenate and coenzyme A biosynthesis III\\|', ''))  %>% 
  # extract the species info
  mutate(pathway = str_extract(pathway, 's__.+$|unclassified'),
         pathway = str_replace(pathway, 's__','')) %>% 
  rename(species = pathway)

# get the top 7 taxa and the unclassfied and other for all the others 
# get the above table 
mean_s <- cnt %>% 
  column_to_rownames('species') %>% 
  as.matrix() %>% 
  apply(., 1, mean) %>% 
  sort(decreasing = T)

# top 8
n <- 8
toppw <- names(mean_s)[1:n]

@@

cleaned_cnt <- bind_rows(splitother %>% 
    pluck('TRUE'),
  splitother %>% 
    pluck('FALSE') %>% 
    gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
    group_by(Sample) %>% 
    summarise(other = sum(relab)) %>% 
    spread(Sample , other) %>% 
    mutate(species = 'Other') %>% 
    select(species, pheno$Sampleid)) %>% 
  # rearrange the species order factor lever
  mutate(species = factor(species, levels = c(toppw[toppw != 'unclassified'] , 'Other', 'unclassified'))) %>% 
  # the long format is the one for plotting
  gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
  left_join(phelabel %>% 
              rename(Sample = Sampleid), by = 'Sample') %>% 
  mutate(Toxicity = factor(Toxicity))

@@

background_col <- data_frame(
  col = bg_col,
  xstart = c(0, phe1_num) + 0.5,
  xend = c(phe1_num, nrow(phelabel))+ 0.5,
  toxicity = names(table(phelabel$Toxicity))
)

# the ggplot
ggplot() +
  geom_bar(data = cleaned_cnt, 
           aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species, color = Toxicity), stat = 'identity') +
  geom_rect(data = background_col, 
            aes(xmin = xstart, xmax = xend, ymin = -Inf, ymax = Inf), 
            fill = background_col$col, 
            alpha = 0.2) +
  geom_bar(data = cleaned_cnt, aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
  #guides(alpha = FALSE) +
  scale_fill_manual(values = hex) +
  labs(y = 'relative abundance',
       x = 'Sample',
       title = str_glue('')) +
  theme_light() +
  theme(
    axis.text.x = element_blank(),
    panel.grid.major = element_blank(), 
    axis.ticks.x = element_blank(),  
    panel.grid.minor = element_blank()
    ) +
  ggsave('../figs/species_pw/PWY-4242: pantothenate and coenzyme A biosynthesis III.jpg', width = 12, height = 8, dpi = 300)
```

## Find how many sig pathways from lefse I wanna plot

```{r}
# load the table that has the unique sig pathways
sigpw <- read_csv('../data/unique_sig_pathway.csv')

# find how many species total it has in all of the lefse sig pathways
siglefse <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv') %>% 
  filter(str_detect(pathway,  paste0(pw$pwid, collapse = '|'))) %>% 
  separate(X1, into = c('pwid','taxa'), sep = '\\|')

sig_num <- siglefse %>% 
  split(f = .$pwid) %>% 
  map_dfc(~ nrow(.)) %>% 
  gather(key = 'pw', value = 'taxa_num') %>% 
  arrange(desc(taxa_num)) %>% 
  separate(pw, into = c('pwid','pw'), sep = ': ')

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
```

@@

# the stratified table is used cuz we wanna know the species info
stratified <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified.tsv') %>% 
  rename(PATHWAYID = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$','')))

@@

## Find how many sig PATHWAYIDs from lefse I wanna plot

```{r}
# load the table that has the unique sig PATHWAYIDs
sigpw <- read_csv('../data/unique_sig_PATHWAYID.csv') %>% 
  # grab the pw id which is gonna be able to match the rownames in the stratified table
  mutate(pwid = str_extract(PATHWAYID, '^.+_PWY|^PWY_\\d{3,4}'))  %>% 
  mutate(pwid = str_replace_all(pwid, '_', '-'))

# find how many species total it has in all of the lefse sig PATHWAYIDs
siglefse <- stratified %>% 
  filter(str_detect(PATHWAYID,  paste0(sigpw$pwid, collapse = '|'))) %>% 
  separate(PATHWAYID, into = c('pwid','taxa'), sep = '\\|')

sig_num <- siglefse %>% 
  split(f = .$pwid) %>% 
  map_dfc(~ nrow(.)) %>% 
  gather(key = 'pw', value = 'taxa_num') %>% 
  arrange(desc(taxa_num)) %>% 
  separate(pw, into = c('pwid','pw'), sep = ': ') %>% 
  # for plotting reason filer on >= 9 rows
  filter(taxa_num >= 9)

## Starting from toxicity

```{r}
species_compo_barplot <- function(PHENOVAR, PATHWAYID) {
  # TWO INPUTS: 
  # The pheno variable: CR, Toxicity, ORR
  # The full PATHWAYID
  # get the pheno lable of the samples

  #PATHWAYID <- 'COA-PWY'
  phelabel <-  pheno %>% 
    select(Sampleid, Toxicity)

  cnt <- stratified %>% 
    filter(str_detect(pathway, PATHWAYID)) %>% 
    # extract the species info
    mutate(pathway = str_extract(pathway, 's__.+$|unclassified'),
           pathway = str_replace(pathway, 's__','')) %>% 
    rename(species = pathway)

  # get the top 7 taxa and the unclassfied and other for all the others 
  # get the above table 
  mean_s <- cnt %>% 
    gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
    group_by(species) %>% 
    summarise(species_sum = sum(relab)) %>% 
    arrange(desc(species_sum)) %>% 
    pull(species)

  # top n (variable) including unclassified
  n <- 8
  toppw <- mean_s[1:n]

@@

  cleaned_cnt <- bind_rows(splitother %>% 
      pluck('TRUE'),
    splitother %>% 
      pluck('FALSE') %>% 
      gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
      group_by(Sample) %>% 
      summarise(other = sum(relab)) %>% 
      spread(Sample , other) %>% 
      mutate(species = 'Other') %>% 
      select(species, pheno$Sampleid)) %>% 
    # rearrange the species order factor lever
    mutate(species = factor(species, levels = c(toppw[toppw != 'unclassified'] , 'Other', 'unclassified'))) %>% 
    # the long format is the one for plotting
    gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)])

  # get the sum of relab per sample cuz we wanna order by this
  sample_relab_sum <- cleaned_cnt %>% 
    group_by(Sample) %>% 
    summarise(sample_sum = sum(relab))

  final_cnt <- cleaned_cnt %>% 
    left_join(sample_relab_sum, by = 'Sample') %>% 
    left_join(phelabel %>% 
                rename(Sample = Sampleid), by = 'Sample') %>% 
    mutate(Toxicity = factor(Toxicity)) %>% 
    arrange(Toxicity, sample_sum)

@@

  background_col <- data_frame(
    col = bg_col,
    xstart = c(0, phe1_num) + 0.5,
    xend = c(phe1_num, nrow(phelabel))+ 0.5
  )

  # the ggplot
  ggplot() +
    geom_bar(data = final_cnt, 
             aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
    geom_rect(data = background_col, 
              aes(xmin = xstart, xmax = xend, ymin = -Inf, ymax = Inf), 
              fill = background_col$col, 
              alpha = 0.2) +
    geom_bar(data = final_cnt, aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
    scale_fill_manual(values = hex ) +
    labs(y = 'relative abundance',
         x = 'Sample',
         title = str_glue('Response to Toxicity NO and YES:\nspecies composition of the pathway abundance\n{PATHWAYID}')) +
    theme_light() +
    theme(
      axis.text.x =  element_text(angle = 40, hjust = 1, size = 8),
      panel.grid.major = element_blank(), 
      axis.ticks.x = element_blank(),  
      panel.grid.minor = element_blank()
      ) +
    ggsave(str_glue('../figs/species_pw/{PATHWAYID}.jpg'), width = 12, height = 8, dpi = 300)
}

sig_num$pwid %>% 
  map(~  species_compo_barplot('Toxicity', .))
